---
title: 'Lab Week 5: time series and a map!'
author: "Julia Dagum"
date: "February 6, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Attach packages 

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
```

## Monthly US energy consumption (renewables)

```{r}
us_renew <- read_csv(here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
```

- make description all lowercase
- only keep observations where the description variable contains the string pattern "consumption"
- remove any observations where the description variable contains the word "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))
```

### Convert 'yyyymm' column to a date

```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# Make a version where I have month and year in separate columns

renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```

### Look at it: 

```{r}
renew_gg <- ggplot(data = renew_date,
                   aes(x = month_sep, 
                       y = value, 
                       group = description)) +
  geom_line(aes(color = description))

renew_gg
```

Updating my colors with paletteer palettes: 

```{r}
renew_gg +
  scale_color_paletteer_d("DresdenColor::bloodrites")
```

### Coerce our renew_parsed to a tsibble (time-series enabled data frame)

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our ts data in a couple different ways: 
```{r}
renew_ts %>% autoplot(value)

# this works because the key was already specified as description above

renew_ts %>% gg_subseries(value)

# renew_ts %>% gg_season(value) - doesn't always work

ggplot(data = renew_parsed, 
       aes(x = month, 
           y = value, 
           group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description, 
             ncol = 1, 
             scales = "free", 
             strip.position = "right") 
```

### Just look at the hydroelectric energy consumption 
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)

ggplot(hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))
```

### What if I want the quarterly average consumption for hydro? 

```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```

### Decompose hydro_ts

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

# window specifies moving avg window
# value as a function of season

components(dcmp) %>% autoplot()

# residual component is usually less than 10% of total range of values (e.g. here it ranges from about 150 - 300 trillion BTUs, errors about 20)
```

```{r}
hist(components(dcmp)$remainder)
```

Now let's look at the ACF: 
```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```
# DANGER DANGER

Make sure to do research on this, don't just do forecasting without a lot of background info

```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

### World map

```{r}
world <- read_sf(dsn = here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"), layer = "TM_WORLD_BORDERS_SIMPL-0.3")
```

